### 2016-06-01 DraftOne.Rmd
First commit (test)

---
title: "DraftOne"
author: "Sarah Koehler"
date: "June 1, 2016"
output: html_document
---

###First draft of eesim.
```{r}
sum(1:4)
```