### 2016-06-23 writing/SimulationPaper/SimulationPaper.Rmd
Add template for simulation paper draft

---
title: Performance of time series, case-crossover, and cross-year models in estimating the effects of natural disasters on human health in the presence of delayed or displaced effects 
author:
  - name: Sarah Koehler
    email: sakoehler7@gmail.com
    affiliation: Colorado State University Statistics
  - name: G. Brooke Anderson
    email: brooke.anderson@colostate.edu
    affiliation: Colorado State University ERHS
    footnote: Corresponding Author
address:
  - code: Colorado State University Statistics
    address: Colorado State University, Department of Statistics, Street, Fort Collins, CO, Zip
  - code: Colorado State University ERHS
    address: Colorado State University, Department of Environmental & Radiological Health Sciences, Lake Street, Fort Collins, CO, Zip
abstract: |
  Time series and case-crossover studies have been extensively used to determine the health effects of both continuous environmental exposures (e.g., air pollution concentration, temperature) and binary exposures (e.g., heat waves). However, exposures to severe natural disasters, like heat waves and tropical storms, could affect mid-term patterns in health risks through either delayed effects or through mortality displacement.

  Here, we conduct a simulation study to determine the performance of time series, case-crossover, and cross-year models to assess the acute health effects of natural disasters under different scenarios of delayed health effects or mortality displacement.

bibliography: mybibfile.bib
output: rticles::elsevier_article
---

Introduction
==========================

Data and Methods
==========================

#### Data

#### Methods

Results
==========================

Discussion
==========================

References {#references .unnumbered}
==========