### 2016-08-01 custom_examples.Rmd
Adding sim_exposure

---
title: "Examples of Customization"
output: pdf_document
---

#Exposure

##calc_t

```{r}
sintrend <- function(n, y){
  day <- c(1:n)
  1 + y * sin(2 * pi * (day / 365))
}
calc_t(n = 5, y = .3, trend = "custom", custom_func = sintrend)
plot(calc_t(1000, y = .3, trend = "custom", custom_func = sintrend))
```

bin_t is similar.

##binary and continuous exposure

```{r}
binary_exposure(n=5, p = .2, y = .1, trend = "custom", custom_func = sintrend)
plot(binary_exposure(n=400, p = .2, y = .1, trend = "custom", custom_func = sintrend))

continuous_exposure(n = 5, y = .6, mu = 100, sd = 10, trend = "custom", custom_func = sintrend)
plot(continuous_exposure(n=1000, y = .6, mu = 5, sd = .5, trend = "custom", custom_func = sintrend))
```