### 2016-08-23 custom_examples.Rmd
Add example custom functions, all options working for sim_outcome

---
title: "Examples of Customization"
output: pdf_document
---

#Exposure

##calc_t

```{r}
sintrend <- function(n, y){
  day <- c(1:n)
  base <- 1 + y * sin(2 * pi * (day / 365))
  rnorm(n, mean = base, sd = 1)
}
calc_t(n = 5, y = .3, trend = "custom", custom_func = sintrend)
plot(calc_t(1000, y = 2, trend = "custom", custom_func = sintrend))
```

bin_t is similar.

##binary and continuous exposure

```{r}
binary_exposure(n=5, p = .2, y = .1, trend = "custom", custom_func = sintrend)
plot(binary_exposure(n=400, p = .2, y = .1, trend = "custom", custom_func = sintrend))

continuous_exposure(n = 5, y = .6, mu = 100, sd = 10, trend = "custom", custom_func = sintrend)
plot(continuous_exposure(n=1000, y = .6, mu = 5, sd = .5, trend = "custom", custom_func = sintrend))
```

##Custom functions for outcome

```{r}
#For baseline:
custombase <- function(n, slope, intercept){
  day <- c(1:n)
  baseline <- day*slope + intercept
  return(baseline)
}

#Example
custombase(n=5, slope = .3, intercept = 55)

#For lambda:
customlambda <- function(exposure, rr, constant, baseline){
  log_lambda <- log(baseline) + log(rr) * exposure + constant
  lambda <- exp(log_lambda)
  return(lambda)
}

#Example:
testlambda <- customlambda(testexp$x, testbase$baseline, rr=1.01, constant = 4)
```